---
title: "R Homework The Analytics Edge Unit 1"
output: html_notebook
---
An analitical detective
https://learning.edx.org/course/course-v1:MITx+15.071x+2T2020/block-v1:MITx+15.071x+2T2020+type@sequential+block@a5915d0492804dada5feb1926ba5be7a/block-v1:MITx+15.071x+2T2020+type@vertical+block@c19c29adcea24760985949691e8019a4
https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2

Problem 1.1 - Loading the Data
How many rows of data (observations) are in this dataset? 191641
Problem 1.2 - Loading the Data
How many variables are in this dataset? 11

```{r}
# Get the current directory
  getwd()
# Read the csv file
  df = read.csv("mvtWeek1.csv")
# Structure of the dataset
  str(df)
# Statistical summary
  summary(df)
```

Problem 1.3 - Loading the Data
Using the "max" function, what is the maximum value of the variable "ID"? 18134
```{r}
which.max(df$ID)
```
Problem 1.4 - Loading the Data
What is the minimum value of the variable "Beat"? 4756
```{r}
which.min(df$Beat)
```
Problem 1.5 - Loading the Data
How many observations have value TRUE in the Arrest variable (this is the number of crimes for which an arrest was made)? 15536
```{r}
table(df$Arrest)
```
Problem 1.6 - Loading the Data
How many observations have a LocationDescription value of ALLEY? 46
```{r}
match("ALLEY", df$LocationDescription)
```

Problem 2.1 - Understanding Dates in R
In many datasets, like this one, you have a date field. Unfortunately, R does not automatically recognize entries that look like dates. We need to use a function in R to extract the date and time. Take a look at the first entry of Date (remember to use square brackets when looking at a certain entry of a variable).

In what format are the entries in the variable Date? Month/Day/Year Hour:Minute
```{r}
df$Date[1]
```

Problem 2.2 - Understanding Dates in R

Now, let's convert these characters into a Date object in R. In your R console, type

DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))

This converts the variable "Date" into a Date object in R. Take a look at the variable DateConvert using the summary function.

What is the month and year of the median date in our dataset? Enter your answer as "Month Year", without the quotes. (Ex: if the answer was 2008-03-28, you would give the answer "March 2008", without the quotes.) May 2006
```{r}
DateConvert = as.Date(strptime(df$Date, "%m/%d/%y %H:%M"))
DateConvert[1]
median(DateConvert, na.rm=TRUE)
```
Problem 2.3 - Understanding Dates in R
Now, let's extract the month and the day of the week, and add these variables to our data frame mvt. We can do this with two simple functions. Type the following commands in R:
```{r}
df$Month = months(DateConvert)

df$Weekday = weekdays(DateConvert)
```
This creates two new variables in our data frame, Month and Weekday, and sets them equal to the month and weekday values that we can extract from the Date object. Lastly, replace the old Date variable with DateConvert by typing:
```{r}
df$Date = DateConvert
```
Using the table command, answer the following questions.
In which month did the fewest motor vehicle thefts occur? Febrero 13511
```{r}
table(df$Month)
```
On which weekday did the most motor vehicle thefts occur? viernes 29284
```{r}
table(df$Weekday)
```
Problem 2.5 - Understanding Dates in R
Which month has the largest number of motor vehicle thefts for which an arrest was made? Enero 1435
```{r}
table(df$Arrest, df$Month)
```
Problem 3.1 - Visualizing Crime Trends
```{r}
hist(df$Date, breaks=100)
```
In general, does it look like crime increases or decreases from 2002 - 2012?
In general, does it look like crime increases or decreases from 2005 - 2008?
In general, does it look like crime increases or decreases from 2009 - 2011?







